Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
setwd('C:/Users/haoli/OneDrive/Documents/GitHub/DataAnalysiswithRUdacity/lesson3')
list.files()
# *.tsv is tab separated values, so the seperator is set to tabs \t.
pf <- read.csv('pseudo_facebook.tsv',sep = '\t')
# read.delim's default separator is \t.
pf2 <- read.delim('pseudo_facebook.tsv')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
# install.packages('ggplot2')
library(ggplot2)

names(pf)
# '+' is used to add additional condition for the bin size
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)
# install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)
theme_set(theme_minimal(24))
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
This first day of each month has an extremely high value compared to other dates, which was not normal. The 31st day of the month has a lower value compared to other dates, which corresponded to that not all months have 31 days.

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: I posted pictures of diving, fishing, and snorkeling while I was in Maldives.

#### How many of your friends do you think saw that post?
Response: 80.

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response: 35

***

### Perceived Audience Size
Notes:
The perceived size as much smaller than the actual size.
***
### Faceting
Notes:
facet_grid(vertical~horizontal) is useful when you want to facet two or more variables.
facet_wrap(~variable) is for one variable.

```{r Faceting}
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month,ncol = 3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:
There were so many people born on the 1st of January. 
That might be because that the default value is Jan 1st, and many people chose to use
it to protect their privacy.

***

### Be Skeptical - Outliers and Anomalies
Notes:
Outliers and anomalies sometimes reporesent bad data, or the limitation of your data.

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response: bad data about an extreme case

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(data = pf, x = friend_count)
```

#### How is this plot similar to Moira's first plot?
Response:
There are outliers with extreme values.
Long-tail data.
***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
# All three plot are equivalent
qplot(data = pf, x = friend_count, xlim = c(0,1000))

qplot(data = pf, x = friend_count) +
  scale_x_continuous(limits = c(0,1000))

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes: Most people guess the audience size at 20, 50, and 100, not numbers in between.
```{r}
qplot(data = pf, x = friend_count, binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0, 1000, 50))
```

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf[pf$gender == "male",], binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
qplot(x = friend_count, data = pf[pf$gender == "female",], binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
qplot(x = friend_count, data = pf,binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_grid(gender ~ .)

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:
Missing values are counted as NA by R.

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)

# na.omit() would remove any observations that have NA in them
qplot(x = friend_count, data = na.omit(pf),binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response:22

#### Why would the median be a better measure than the mean?
Response:This is a skewed distribution.

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure, data = pf, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x = tenure/365, data = pf, binwidth = 0.25,
      color = I('black'), fill = I('#099DD9'))+
  scale_x_continuous(limits = c(0,7), breaks = seq(1,7,1))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x = tenure/365, data = pf, binwidth = 0.25, 
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420'))+
  scale_x_continuous(limits = c(0,7), breaks = seq(1,7,1))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = pf, binwidth = 1,
      xlab = 'User Age',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#5760AB')) +
  scale_x_continuous(breaks = seq(0,113,5))

qplot(x = mobile_likes, data = pf,binwidth = 10,
      xlab = 'Number of Mobile Likes',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#5760AB')) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(100,1000,100)) +
  scale_y_continuous(limits = c(0, 1100))

```

#### What do you notice?
Response:
There are peaks in the age of 20, 25, and 27. Those peaks greater 100 might be people exaggerating their ages.

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes: They tended to reoccur sometimes.


***

### Transforming Data
Notes:
engaged values means values that have a long tail.
Data is over dispersed.
```{r}
qplot(x=friend_count, data = pf)

summary(pf$friend_count)
summary(log10(pf$friend_count + 1))
summary(sqrt(pf$friend_count))

# install.packages('gridExtra')
library(gridExtra)

p1 = qplot(x = friend_count, data = pf)
p2 = qplot(x = log10(friend_count + 1), data = pf,
           xlab = 'Number of friends',
           ylab = 'log base 10 of number of users in the sample')
p3 = qplot(x = sqrt(friend_count), data = pf,
           xlab = 'Number of friends',
           ylab = 'square root of number of users in the sample')
grid.arrange(p1, p2, p3, ncol=1)

```

***
### Transformating Data Alternate Solution
### Use scales!
Notes:

```{r}
logScale <- qplot(x = log10(friend_count), data = pf)

countScale <- ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_log10()

grid.arrange(logScale, countScale, ncol=2)
```

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol=1)
```

***


### Frequency Polygons (before we had histograms, plot two plots on top of each other)

```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)

qplot(x = friend_count, y = ..count../sum(..count..),
        data = subset(pf, !is.na(gender)),
      xlab = 'Friend Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

***
### Frequency Polygons Solution
```{r}
qplot(x = www_likes, y = ..count../sum(..count..), data = subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender)
# instructor's solution
qplot(x = www_likes, y = ..count../sum(..count..), data = subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender) +
  scale_x_log10()
# my solution, not very good  
qplot(x = www_likes, y = ..count../sum(..count..), data = subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender, binwidth = 1,
      xlab = 'www like count',
      ylab = 'Proportion of Users with that www like count') +
  scale_x_continuous(limits = c(0,10), breaks = c(0,10,1))
```

### Likes on the Web
Notes:

```{r Likes on the Web}
by(pf$www_likes,pf$gender, sum)

```


***
### Histograms
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot')
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
# This way, we removed data from calculation
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0,1000))

# Using coord_cartesian()
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 250))
by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response: women
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
names(pf)
qplot(x = gender, y = friendships_initiated, data = subset(pf,!is.na(gender)), geom = 'boxplot') +
  coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:
I used by() function to compare the median number of friends count for different genders, women and men, and then figured out that women make more friends than men on average.
***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)
summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_check_in > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
63947/99003

sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:
Understanding the details of one variable.
In this lesson, I learnt how to make histogram, boxplot, and frequency polygons for one variable of interest, based on different categories if necessary, to know the basic statistics of the data. I learnt to make some adjustments when analyzing data, in order to exclude the outliers, choose an appropriate data limit, or adjust the bin.
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!