---
title: 'Lesson 6 Problem Set: Explore Two Variables'
author: "MZ"
date: "February 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. In this problem set, you'll continue
# to explore the diamonds data set.

# Your first task is to create a
# scatterplot of price vs x.
# using the ggplot syntax.

```{r diamonds}
library(ggplot2)
diamonds <- diamonds
ggplot(aes(x = x, y = price), data = diamonds) +
  geom_point()
```

# 2. What are your observations about the scatterplot of price vs. x?

The price of the diamonds increases as the x value of the diamonds increases, and the it creases faster and faster. There are some points where the x value is around zero, but the price goes up high, up tp $18000.

# 3. Correlation
# What is the correlation between price and x?
# What is the correlation between price and y?
# What is the correlation between price and z?

```{r}
with(diamonds, cor.test(x, price))
with(diamonds, cor.test(y, price))
with(diamonds, cor.test(z, price))
```

# 4. Create a simple scatter plot of price vs depth.

```{r scatter plot}
ggplot(aes(x = depth, y = price), data = diamonds) +
  geom_point()
```

# 5. Change the code to make the transparency of the points to be 1/100 of what they are now and mark the x-axis every 2 units. See the instructor notes for two hints.

Hint 1: Use the alpha parameter in geom_point() to adjust the transparency of points. 
Hint 2: Use scale_x_continuous() with the breaks parameter to adjust the x-axis.
```{r}

ggplot(aes(x = depth, y = price), data = diamonds) +
  geom_point(alpha = 1/100) +
  scale_x_continuous(breaks = seq(0, 80, 2))
```
# 6. Based on the scatter plot of depth vs. price, most diamonds are between what values of depth?
58 an 64

# 7. What is the correlation of depth vs. price?
```{r}
with(diamonds, cor.test(depth, price))
```

# Based on the correlation coefficient, would you use depth to predict the price of a diamond? Why?
No. The correlation between diamond depth and price is too weak.

# 8. Create a scatterplot of price vs carat and omit the top 1% of price and carat values.
```{r}
price_99 <- quantile(diamonds$price, .99)
carat_99 <- quantile(diamonds$carat, .99)
ggplot(aes(carat, price), data = subset(diamonds, price <= price_99 | carat <= carat_99)) +
  geom_point(alpha = 1/20)
  
```

# 9. Create a scatterplot of price vs. volume (x * y * z). This is a very rough approximation for a diamond's volume.

# Create a new variable for volume in the diamonds data frame.
# This will be useful in a later exercise.

# Don't make any adjustments to the plot just yet.
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z
ggplot(aes(volume, price), data = diamonds) +
  geom_point()
```

# 10. Waht are your observations from the price vs. volume scatter plot?
The price of diamond increases as the volume increases, and increases more rapidly. The volume of the diamond is mostly below 500 mm^3.
Did you notice some outliers? Some volumes are 0! There's an expensive diamond with a volume near 4000, and a cheaper diamond with a volume near 900.

# You can find out how many diamonds have 0 volume by using count(diamonds$volume == 0). The count() function comes with the plyr package.

# Note: If you ran the count function from plyr, you need to run this command in R to unload the plyr package.
## detach("package:plyr", unload=TRUE)
# The plyr package will conflict with the dplyr package in later exercises.

# 11. What's the correlation of price and volume? Exclude diamonds that have a volume of 0 or that are greater than or equal to 800.

Hint 1: It's helpful to add a variable for volume to the diamonds data frame. You should exclude diamonds that have a volume greater than or equal to 800. Also, exclude diamonds that have a volume of 0. Use the & symbol between two conditions when you subset the original data frame. 
Hint 2: Subset the diamonds data frame under the conditions and save it into another variable. Use that data frame in the cor() function.

```{r}
with(subset(diamonds, volume != 0 & volume < 800), cor.test(volume, price))
```

# 12. Subset the data to exclude diamonds with a volume greater than or equal to 800. Also, exclude diamonds with a volume of 0. Adjust the transparency of the points and add a linear model to the plot. (See the Instructor Notes or look up the documentation of geom_smooth() for more details about smoothers.)

# We encourage you to think about this next question and to post your thoughts in the discussion section.
# Do you think this would be a useful model to estimate the price of diamonds? Why or why not?
I think so. Since the smooth line matches in the middle of the data points.

```{r}
ggplot(aes(volume, price), data = subset(diamonds, volume != 0 & volume < 800)) +
  geom_point(alpha = 1/10) +
  geom_smooth(method = 'lm', color = 'red')
```

# 13. Use the function dplyr package to create a new data frame containing info on diamonds by clarity.

# Name the data frame diamondsByClarity

# The data frame should contain the following variables in this order.

#       (1) mean_price
#       (2) median_price
#       (3) min_price
#       (4) max_price
#       (5) n

# where n is the number of diamonds in each level of clarity.

Note: If you used the count() function from the plyr package before this exercise, you need to run this command to unload the plyr package: detach("package:plyr", unload=TRUE)
The plyr package will conflict with the dplyr package when doing this exercise. You will want to complete this exercise in RStudio with ONLY the dplyr package loaded.

The dplyr package.
Another warning: Version 0.4.0 of dplyr has a bug when using the median function on the summarize layer, depending on the nature of the data being summarized. You may need to cast the data as a numeric (float) type to get the expected results, e.g. median(as.numeric(var)).

```{r}
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n()) %>%
  arrange(clarity)
```

# 14. We've created summary data frames with the mean price by clarity and color. You can run the code in R to verify what data is in the variables diamonds_mp_by_clarity and diamonds_mp_by_color.

# Your task is to write additional code to create two bar plots on one output image using the grid.arrange() function from the package gridExtra.

Hint: Save each plot into a variable and pass the two variables to grid.arrange(). Be sure you have installed and loaded the gridExtra package.

```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(aes(clarity, mean_price), data = diamonds_mp_by_clarity) +
  geom_col()
p2 <- ggplot(aes(color, mean_price), data = diamonds_mp_by_color) +
  geom_col()

library(gridExtra)
grid.arrange(p1, p2, ncol = 2)

```

# 15. What do you notice in each of the bar chars for mean price by clarity and mean price by color.
The mean price decreases as the clarity of the diamond gets better, and there is a peak mean price at the clarity level of SI1. 
The mean price increases as the color of the diamond gets worse, but the D level color has a slightly higher mean price than the E level color diamonds.

We think something odd is going here. These trends seem to go against our intuition.
Mean price tends to decrease as clarity improves. The same can be said for color.
We encourage you to look into the mean price across cut.

```{r}
diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

ggplot(aes(cut, mean_price), data = diamonds_mp_by_cut) +
  geom_col()
```


# 16. The Gapminder website contains over 500 data sets with information about the world's population. Your task is to continue the investigation you did at the end of Problem Set 3 or you can start fresh and choose a different data set from Gapminder.

# If you're feeling adventurous or want to try some data munging see if you can find a data set or scrape one from the web.

# In your investigation, examine pairs of variable and create 2-5 plots that make use of the techniques from Lesson 4.

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
# 1. the variable(s) you investigated, your observations, and any summary statistics
# 2. snippets of code that created the plots
# 3. links to the images of your plots

Gapminder Data
https://www.gapminder.org/data/
Hans Rosling's 200 Countries, 200 Years, 4 Minutes
https://www.youtube.com/watch?v=jbkSRLYSojo
Save plots using ggsave()

```{r}
BMI_female <- read.csv("C:/Users/mengn/OneDrive/Documents/GitHub/DataAnalysiswithRUdacity/lesson3/indicator_BMI female ASM.csv", check.names = F, 1)
SBP_female <- read.csv("C:/Users/mengn/OneDrive/Documents/GitHub/DataAnalysiswithRUdacity/lesson3/Indicator_SBP female ASM.csv", check.names=F, 1)
TC_female <- read.csv("C:/Users/mengn/OneDrive/Documents/GitHub/DataAnalysiswithRUdacity/lesson3/Indicator_TC female ASM.csv", check.names=FALSE)
Breast_Death_female <- read.csv("C:/Users/mengn/OneDrive/Documents/GitHub/DataAnalysiswithRUdacity/lesson3/indicator breast female mortality.csv", check.names=FALSE)
#colnames(BMI_female)[1] <- 'country'
#colnames(SBP_female)[1] <- 'country'
library(tidyr)
BMI <- gather(BMI_female, "year","BMI", 2:length(colnames(BMI_female)))
names(BMI)[[1]] <- "country"
SBP <- gather(SBP_female, "year","SBP", 2:length(colnames(SBP_female)))
names(SBP)[[1]] <- "country"

BMI_and_SBP <- inner_join(BMI,SBP, by = c('country','year'))
ggplot(aes(BMI, SBP), data = BMI_and_SBP) +
  geom_point(alpha = 1/10)+
  labs(title = 'Female SBP vs. BMI')

BMI_and_SBP_comb <- subset(BMI_and_SBP, country == 'China' | country == 'United States'|country == 'Russia' | country == 'Korea, Dem. Rep.')
ggplot(aes(BMI, SBP), data = BMI_and_SBP_comb) +
  geom_point()+
  facet_wrap(~country)

ggplot(aes(BMI, SBP), data = BMI_and_SBP_comb) +
  geom_smooth()+
  facet_wrap(~country)

ggplot(aes(BMI, SBP), data = BMI_and_SBP_comb) +
  geom_line()+
  facet_wrap(~country)
```

