Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x=age, y=friend_count, data=pf)
qplot(age, friend_count, data=pf)
```

***

#### What are some things that you notice right away?
Response:
The plot is binomial distributed, with a lot of people at around age 1-3 with a huge count of thousands of friends.
There are some vertical bars indicating people lying about their age, at 69 and 100.

***

### ggplot Syntax
Notes:
ggplot is more formal. It will help us with more complicated plots.

```{r ggplot Syntax}
qplot(x=age, y=friend_count, data=pf)

ggplot(aes(x=age, y=friend_count), data = pf) +
  geom_point() +
  xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes:
Makes it difficult to see all the dots

```{r Overplotting}
# alpha for transparency, 20 of the dots to be equivalent to one block dot.
# geom_jitter adds some noise to each age, make it more dispersed
ggplot(aes(x=age, y=friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:
The data points seem to spread out smoothly along the age, and from the darkness of the dot clusters, we can see that the bulk of young users have friends count of less than 1000 for the youth, and less than 500 for age over 30.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x=age, y=friend_count), data = pf) +
  geom_point(alpha=1/20) +
  xlim(13, 90) 
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

jitter gives us noise, so could gives us friend count below zero, which is not feasible for squre root.
We need to set the position parameter equal to position jitter, with a mininum height of 0.
```{r}
ggplot(aes(x=age, y=friend_count), data = pf) +
  geom_point(alpha=1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) +
  coord_trans(y = "sqrt")
```

#### What do you notice?
The bulk friend count for young users is below 1000.
***

### Alpha and Jitter
Notes:
Explore the relationship between friends initiated vs age.

```{r Alpha and Jitter}
summary(pf$age)
ggplot(aes(age, friendships_initiated), data = pf) +
  geom_point(alpha = 1/10, position = position_jitter(h = 0)) +
  xlim(13,90) +
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages("dplyr")
library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```
```

Create your plot!

```{r Conditional Means Plot}
library(ggplot2)
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age)+
geom_line()+
xlim(13,90)
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
summary(pf$age)
ggplot(aes(age, friendships_initiated), data = pf) +
  geom_point(alpha = 1/10, position = position_jitter(h = 0), 
  color = 'orange') +
  coord_cartesian(xlim = c(13,90), ylim = c(0, 3000)) +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1),
  linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9),
  linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .5),
  color = 'blue')
```

#### What are some of your observations of the plot?
Response:
Even for the yound users, it's very unlikely to have friends over 1000.
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = "pearson")
with(pf, cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf,age<=70) , cor.test(age,friend_count))
```

***

### Correlation Methods
Notes:

***
Other than pearson method, we can also use kendall or spearman methods.

## Create Scatterplots
Notes:
```{r}
library(ggplot2)
ggplot(aes(y = likes_received, x = www_likes_received), data = pf, 
alpha = 1/20) +
geom_point() +
scale_x_log10()+
scale_y_log10()
```


***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(y = likes_received, x = www_likes_received), data = pf) +
geom_point() +
xlim(0, quantile(pf$www_likes_received, 0.95)) +
ylim(0, quantile(pf$likes_received, 0.95)) +
geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf,cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(Month, Temp), data = Mitchell) +
geom_point()

ggplot(aes(Month, Temp), data = Mitchell) +
geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0.002
b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Month, Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(Month, Temp), data = Mitchell) +
geom_point() +
scale_x_continuous(breaks = seq(0, 203, 12))

ggplot(aes(Month%%12, Temp), data = Mitchell) +
geom_point()
```

***

### A New Perspective

What do you notice?
Response:
Temperatures are higher in the middle of the year, and lower at the beginning and ending of the year. The temperature in each year tends to cluster within certain range.

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x=age, y= friend_count_mean), data = pf.fc_by_age) +
geom_line()

head(pf.fc_by_age, 10)
pf.fc_by_age[17:19,]
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf$age + (1 - pf$dob_month/12)
```

Programming Assignment
```{r Programming Assignment}

library(dplyr)
pf.fc_by_age_months <- pf %>%
group_by(age_with_months) %>%
summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>%
arrange(age_with_months)
head(pf.fc_by_age_months)

age_with_months_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months2 <- summarise(age_with_months_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age_months2 <- arrange(pf.fc_by_age_months2, age_with_months)
head(pf.fc_by_age_months2)


```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean),
data = subset(pf.fc_by_age, age <71)) +
geom_line() +
geom_smooth()

p2 <- ggplot(aes(age_with_months, friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) +
geom_line() +
geom_smooth()

p3 <- ggplot(aes(x = round(age/5)*5, y = friend_count), data = subset(pf, age < 71))+
geom_line(stat = 'summary', fun.y = mean)
library(gridExtra)
grid.arrange(p2,p1,p3, ncol = 1)
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

